---
title: "PAWN Sensitivity Analysis"
author: "Sumit Jyoti"
date: "2026-08-22"
output:
  html_document:
    toc: true
    toc_depth: 3
    number_sections: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
rm(list = ls())
source("model_functions.R")
```

# Model parameters

```{r model-parameters}
alpha <- 0.60; beta <- 1.31e6; tfs <- 1; MIC <- 0.13
fr_min <- 0; fr_max <- 0.5; pMIC_min <- 0; pMIC_max <- 10
C_min <- 75; C_max <- 1100
M_shape <- 0.8760928; M_scale <- 43.276368
F_min <- 0.0018; F_max <- 0.0233

n <- 15; Nu <- 100; Nc <- 100
pawn_seed <- 0; ncores <- 4
critval <- 1.36 * sqrt((Nu + Nc) / (Nu * Nc))
```

# Pharmacodynamic parameters

$E_{max}$ and $EC_{50}$ are sampled independently so that their effects can be
evaluated separately by PAWN. Positive values are retained, and $E_{max}$ is
converted from log10 per hour to natural-log per day for the SD-DRM.

```{r pd-parameters}
Emax_mean <- 1.476379; Emax_sd <- sqrt(0.01068527)
EC50_mean <- 0.1668759; EC50_sd <- sqrt(0.00210130)

rposnorm <- function(N, mean, sd) {
  x <- rnorm(N, mean, sd)
  while (any(x <= 0)) x[x <= 0] <- rnorm(sum(x <= 0), mean, sd)
  x
}

sample_Emax <- function(N) rposnorm(N, Emax_mean, Emax_sd) * 24 * log(10)
sample_EC50 <- function(N) rposnorm(N, EC50_mean, EC50_sd)
```

# PAWN functions

PAWN measures sensitivity using the maximum distance between the unconditional
output distribution and output distributions obtained by conditioning each
input. The SD-DRM lookup table is generated once from the sampled
pharmacodynamic range and reused for all model evaluations.

```{r pawn-functions}
evaluate_model <- function(data, model, ab_lookup, ncores = 1) {
  f <- function(i) model(data[i, , drop = FALSE], ab_lookup)

  if (ncores > 1 && .Platform$OS.type != "windows") {
    unlist(parallel::mclapply(
      seq_len(nrow(data)), f,
      mc.cores = ncores,
      mc.set.seed = FALSE
    ))
  } else {
    vapply(seq_len(nrow(data)), f, numeric(1))
  }
}

PAWN <- function(model, input_names, sample_inputs, Nu = 100, n = 15,
                 Nc = 100, seed = 0, ncores = 1) {
  set.seed(seed)
  M <- length(input_names)
  par_u <- sample_inputs(Nu)
  par_c <- sample_inputs(M * n * Nc)

  for (j in seq_len(M)) {
    for (i in seq_len(n)) {
      idx <- ((j - 1) * n * Nc + (i - 1) * Nc + 1):
        ((j - 1) * n * Nc + i * Nc)

      par_c[idx, input_names[j]] <-
        sample_inputs(1)[[input_names[j]]][1]
    }
  }

  all_inputs <- rbind(par_u, par_c)

  ab_lookup <- make_ab_lookup(
    alpha, beta,
    all_inputs$Emax,
    all_inputs$EC50,
    max(all_inputs$pMIC) / 100 * MIC,
    tfs,
    n_grid = 250,
    seed = seed,
    nsim = 1000
  )

  y_u <- evaluate_model(par_u, model, ab_lookup, ncores)
  y_c <- evaluate_model(par_c, model, ab_lookup, ncores)
  KS <- matrix(NA_real_, M, n, dimnames = list(input_names, NULL))

  for (j in seq_len(M)) {
    for (i in seq_len(n)) {
      idx <- ((j - 1) * n * Nc + (i - 1) * Nc + 1):
        ((j - 1) * n * Nc + i * Nc)

      KS[j, i] <- unname(suppressWarnings(
        ks.test(y_u, y_c[idx], exact = FALSE)$statistic
      ))
    }
  }

  list(KS = KS)
}
```

# SD-DRM sensitivity model

Retail concentration, serving mass, and pathogenic fraction represent
variability only. The resulting PAWN indices therefore describe the influence
of the complete exposure-input distributions.

```{r sensitivity-model}
rlogunif <- function(N, min, max) {
  exp(runif(N, log(min), log(max)))
}

sample_inputs <- function(N) {
  data.frame(
    C_retail = rlogunif(N, C_min, C_max),
    M = rgamma(N, shape = M_shape, scale = M_scale),
    F_pathogenic = runif(N, F_min, F_max),
    fr = runif(N, fr_min, fr_max),
    pMIC = runif(N, pMIC_min, pMIC_max),
    Emax = sample_Emax(N),
    EC50 = sample_EC50(N)
  )
}

risk_model <- function(x, ab_lookup) {
  dose <- x$C_retail * x$M * x$F_pathogenic

  calc_sd_drm(
    dose,
    x$fr,
    x$pMIC / 100 * MIC,
    alpha,
    beta,
    x$Emax,
    x$EC50,
    tfs,
    ab_lookup
  )$risk_total[1]
}

input_names <- c(
  "C_retail", "M", "F_pathogenic",
  "fr", "pMIC", "Emax", "EC50"
)

res <- PAWN(
  risk_model,
  input_names,
  sample_inputs,
  Nu,
  n,
  Nc,
  pawn_seed,
  ncores
)
```

# Sensitivity results and figure

The PAWN index is the Kolmogorov-Smirnov distance between unconditional and
conditional output distributions. Higher values indicate a stronger influence
on illness risk. The dashed line is a descriptive two-sample KS reference
threshold, not a formal classification rule.

```{r pawn-summary}
median_pawn <- apply(res$KS, 1, median)

summary_pawn <- data.frame(
  Parameter = c(
    "C_retail", "M", "F_pathogenic",
    "f_r", "pMIC", "Emax", "EC50"
  ),
  Rank = rank(-median_pawn, ties.method = "min"),
  Median_PAWN = median_pawn,
  Min_PAWN = apply(res$KS, 1, min),
  Max_PAWN = apply(res$KS, 1, max),
  Median_vs_threshold = ifelse(
    median_pawn > critval,
    "Median above threshold",
    "Median below threshold"
  )
)

summary_pawn <- summary_pawn[order(summary_pawn$Rank), ]

knitr::kable(
  summary_pawn,
  digits = 3,
  caption = "PAWN sensitivity indices"
)
```

```{r pawn-figure, fig.width=8.5, fig.height=5, dpi=600, fig.path="images/PAWN_"}
plot_pawn <- function(res, labels) {
  ord <- order(apply(res$KS, 1, median), decreasing = TRUE)
  KS <- res$KS[ord, , drop = FALSE]
  labels <- labels[ord]

  boxplot(
    t(KS),
    ylim = c(0, 1),
    outline = FALSE,
    xaxt = "n",
    xlab = "",
    ylab = "PAWN index",
    cex.axis = 1.1,
    cex.lab = 1.2
  )

  axis(1, at = seq_along(labels), labels = FALSE)

  text(
    seq_along(labels),
    -0.09,
    labels,
    srt = 40,
    adj = 1,
    xpd = TRUE,
    cex = 0.95
  )

  points(
    jitter(rep(seq_along(labels), n), factor = 1.5),
    as.vector(KS),
    pch = 21,
    bg = "grey70"
  )

  abline(h = critval, lty = 2)
}

labels <- expression(
  C[retail], M, F[pathogenic],
  f[r], "pMIC", E[max], EC[50]
)

set.seed(pawn_seed)
par(mar = c(5.5, 4.2, 1, 1))
plot_pawn(res, labels)
```

# Analysis settings

```{r analysis-settings}
settings <- data.frame(
  Setting = c(
    "Conditioning values",
    "Unconditional samples",
    "Conditional samples",
    "KS reference value",
    "Exposure inputs",
    "Emax-EC50 dependence"
  ),
  Value = c(
    n,
    Nu,
    Nc,
    round(critval, 4),
    paste(
      "C_retail, M, and F_pathogenic:",
      "variability only"
    ),
    "Independent"
  )
)

knitr::kable(
  settings,
  caption = "PAWN sensitivity analysis settings"
)
```
