---
title: "SD-DRM Analysis"
author: "Sumit Jyoti"
date: "2026-08-22"
output:
  html_document:
    toc: true
    toc_depth: 3
    number_sections: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
rm(list = ls())

library(dplyr)
library(tidyr)
library(purrr)
library(ggplot2)

source("model_functions.R")

seed <- 123
set.seed(seed)
```

# Exposure data and model parameters

The exposure file contains the two-dimensional simulation from the exposure
assessment. Each column of the dose matrix represents one uncertainty draw,
and each row represents variability among consumption events within that draw.

```{r data-parameters}
exposure <- read.csv("raw_shrimp_mc2d_simulation_data.csv") |>
  arrange(inner_id)

n_inner <- nrow(exposure)
n_outer <- 500 

dose_inner <- exposure$D_raw
dose <- rep(dose_inner, times = n_outer)

alpha <- 0.60
beta <- 1.31e6
tfs <- 1
MIC <- 0.13

fr_fixed <- 0.20
pMIC_fixed <- 2

pMIC_seq <- seq(0, 10, by = 0.1)
pMIC_plot <- c(0, 2, 5)
pMIC_breaks <- 0:10

fr_seq <- seq(0, 0.5, by = 0.025)
fr_plot <- c(0, 0.3, 0.5)
fr_breaks <- seq(0, 0.5, by = 0.05)


pMIC_colours <- c("0% MIC" = "#F8766D", "2% MIC" = "#00BFC4", "5% MIC" = "#C77CFF")
fr_colours <- c("fr = 0" = "#F8766D", "fr = 0.3" = "#7CAE00", "fr = 0.5" = "#C77CFF")
```

# Pharmacodynamic uncertainty

The Emax and EC50 estimates are sampled together from their fitted bivariate
normal distribution. Sampling them as pairs preserves their estimated
relationship. Emax is converted from log10 per hour to natural-log per day,
which is the scale used by the SD-DRM.

```{r pd-uncertainty}
pd_estimates <- c(Emax_log10_h = 1.476380, EC50_mg_L = 0.166877)

pd_covariance <- matrix(c(0.010685300, 0.002246552, 0.002246552, 0.002101326), nrow = 2, byrow = TRUE)

draw_pd <- function(n) {
  x <- MASS::mvrnorm(n, pd_estimates, pd_covariance)
  invalid <- x[, 1] <= 0 | x[, 2] <= 0

  while (any(invalid)) {
    x[invalid, ] <- MASS::mvrnorm(sum(invalid), pd_estimates, pd_covariance)
    invalid <- x[, 1] <= 0 | x[, 2] <= 0
  }

  data.frame(
    Emax_per_day = x[, 1] * 24 * log(10),
    EC50_ug_mL = x[, 2]
  )
}

set.seed(seed)
pd <- draw_pd(n_outer)
Emax_outer <- pd$Emax_per_day
EC50_outer <- pd$EC50_ug_mL
```

# SD-DRM simulation

The original beta-Poisson alpha and beta describe the response without an
antibiotic effect. They can therefore be used directly for the resistant
subpopulation, but not for susceptible bacteria exposed to doxycycline.
`make_ab_lookup()` generates r values from the original beta-Poisson model,
adjusts them using the doxycycline concentration and paired Emax-EC50 draws,
and obtains the adjusted alpha and beta for the susceptible subpopulation.
`calc_sd_drm()` then splits the dose into susceptible and resistant components
and combines their extinction probabilities to calculate the total illness
risk.

```{r scenario-functions}
ab_lookup <- make_ab_lookup(
  alpha = alpha,
  beta = beta,
  Emax = Emax_outer,
  EC50 = EC50_outer,
  C_max = max(pMIC_seq) / 100 * MIC,
  tfs = tfs,
  n_grid = 250,
  seed = 0,
  nsim = 1000
)

run_scenario <- function(fr, pMIC) {
  out <- calc_sd_drm(
    dose = dose,
    fr = fr,
    C = pMIC / 100 * MIC,
    alpha = alpha,
    beta = beta,
    Emax = rep(Emax_outer, each = n_inner),
    EC50 = rep(EC50_outer, each = n_inner),
    tfs = tfs,
    ab_lookup = ab_lookup
  )

  list(
    risk = matrix(out$risk_total, nrow = n_inner, ncol = n_outer),
    delta = matrix(out$delta_p_treat, nrow = n_inner, ncol = n_outer)
  )
}

summarise_risk <- function(vary_by, values, fixed_fr = NULL, fixed_pMIC = NULL) {
  map_dfr(values, function(value) {
    fr <- if (vary_by == "fr") value else fixed_fr
    pMIC <- if (vary_by == "pMIC") value else fixed_pMIC
    x <- run_scenario(fr, pMIC)

    q <- quantile(colMeans(x$risk), c(0.025, 0.50, 0.975), names = FALSE)
    prop_more <- mean(x$delta > 0)

    data.frame(
      value = value,
      lower95_expected_risk = q[1],
      median_expected_risk = q[2],
      upper95_expected_risk = q[3],
      prop_more_likely_treatable = prop_more,
      outcome = ifelse(prop_more > 0.5,
                       "More likely treatable", "Less likely treatable")
    )
  })
}

find_cutoff <- function(data, vary_by, fixed_fr = NULL, fixed_pMIC = NULL,
                        tolerance = 0.0001) {
  data <- arrange(data, value)
  above <- data$prop_more_likely_treatable > 0.5
  i <- which(head(above, -1) & !tail(above, -1))[1]
  if (is.na(i)) return(NA_real_)

  uniroot(
    function(value) {
      summarise_risk(vary_by, value, fixed_fr, fixed_pMIC)$prop_more_likely_treatable - 0.5
    },
    interval = data$value[c(i, i + 1)],
    tol = tolerance
  )$root
}
```

# Conditional dose-response curves

These curves use a fixed dose grid, so they show dose-response uncertainty from
the paired Emax-EC50 draws without exposure variability. Dose and illness
probability are transformed to log10 for plotting, while the SD-DRM calculations
remain on their original scales.

```{r dose-response-functions}
dose_grid <- 10^seq(log10(min(dose)), log10(max(dose)), length.out = 300)

point_id <- unique(round(seq(1, length(dose_grid), length.out = 10)))

run_dose_grid <- function(fr, pMIC) {
  out <- calc_sd_drm(
    dose = rep(dose_grid, each = n_outer),
    fr = fr,
    C = pMIC / 100 * MIC,
    alpha = alpha,
    beta = beta,
    Emax = rep(Emax_outer, times = length(dose_grid)),
    EC50 = rep(EC50_outer, times = length(dose_grid)),
    tfs = tfs,
    ab_lookup = ab_lookup
  )

  list(
    risk = matrix(out$risk_total, nrow = n_outer),
    delta = matrix(out$delta_p_treat, nrow = n_outer),
    status = matrix(out$status, nrow = n_outer)
  )
}

make_dose_response <- function(vary_by, values, panel_values) {
  expand_grid(panel_value = panel_values, value = values) |>
    pmap_dfr(function(panel_value, value) {
      if (vary_by == "pMIC") {
        pMIC <- value
        fr <- panel_value
        panel <- paste0("Resistant fraction = ", panel_value)
        group <- paste0(value, "% MIC")
      } else {
        fr <- value
        pMIC <- panel_value
        panel <- paste0("pMIC = ", panel_value, "% MIC")
        group <- paste0("fr = ", value)
      }

      out <- run_dose_grid(fr, pMIC)
      q <- apply(out$risk, 2, quantile, probs = c(0.025, 0.50, 0.975), names = FALSE)

      data.frame(
        dose = dose_grid,
        lower95 = q[1, ],
        median_risk = q[2, ],
        upper95 = q[3, ],
        group = group,
        panel = panel,
        status = apply(out$status, 2, function(x) names(which.max(table(x)))),
        show_point = seq_along(dose_grid) %in% point_id
      )
    })
}

make_pairwise_data <- function(data, comparisons) {
  imap_dfr(comparisons, function(groups, comparison_name) {
    filter(data, group %in% groups) |>
      mutate(comparison = comparison_name)
  })
}

save_plot <- function(plot, file, width, height) {
  dir.create("images", showWarnings = FALSE)
  print(plot)
  ggsave(file.path("images", file), plot, width = width, height = height, dpi = 600, bg = "white")
}

plot_dose_response_2d <- function(
    data, title, group_title, colours, file_name
) {
  stopifnot(all(data$dose > 0, data$lower95 > 0,
                data$median_risk > 0, data$upper95 > 0))

  data <- mutate(
    data,
    log10_dose = log10(dose),
    log10_lower95 = log10(lower95),
    log10_median_risk = log10(median_risk),
    log10_upper95 = log10(upper95)
  )

  p <- ggplot(data, aes(log10_dose, colour = group, fill = group)) +
    geom_ribbon(
      aes(ymin = log10_lower95, ymax = log10_upper95),
      alpha = 0.45, colour = NA, show.legend = FALSE
    ) +
    geom_line(aes(y = log10_median_risk), linewidth = 1.05) +
    geom_point(
      data = filter(data, show_point),
      aes(y = log10_median_risk, shape = status),
      size = 2
    ) +
    facet_grid(comparison ~ panel) +
    scale_x_continuous(breaks = scales::breaks_pretty(n = 6)) +
    scale_y_continuous(breaks = scales::breaks_pretty(n = 6)) +
    scale_colour_manual(
      name = group_title,
      values = colours,
      breaks = names(colours)
    ) +
    scale_fill_manual(
      name = group_title,
      values = colours,
      breaks = names(colours)
    ) +
    scale_shape_manual(
      name = "Treatment outcome",
      values = c(
        "Less likely treatable" = 17,
        "More likely treatable" = 16
      )
    ) +
    labs(
      title = title,
      x = expression(log[10] * "(pathogenic dose [MPN per consumption event])"),
      y = expression(log[10] * "(" * P[illness] * ")"),
      caption = expression(
        "Shaded areas show 95% uncertainty intervals from paired " *
          E[max] * "-" * EC[50] * " draws."
      )
    ) +
    guides(fill = "none") +
    theme_bw(base_size = 11) +
    theme(
      plot.title = element_text(face = "bold"),
      strip.text = element_text(face = "bold"),
      panel.grid.minor = element_blank(),
      legend.position = "top",
      legend.box = "horizontal"
    )

  save_plot(p, file_name, width = 12, height = 9)
}
```

## Residual doxycycline concentration

```{r dose-response-pmic}
dose_response_pMIC <- make_dose_response("pMIC", pMIC_plot, panel_values = c(0, fr_fixed))

pMIC_comparisons <- make_pairwise_data(
  dose_response_pMIC,
  list(
    "0% MIC versus 2% MIC" = c("0% MIC", "2% MIC"),
    "0% MIC versus 5% MIC" = c("0% MIC", "5% MIC")
  )
)

plot_dose_response_2d(
  pMIC_comparisons,
  title = "Dose-response curves by residual doxycycline concentration",
  group_title = "Residual doxycycline concentration (% MIC)",
  colours = pMIC_colours,
  file_name = "dose_response_varying_pMIC.png"
)
```

## Resistant fraction

```{r dose-response-fr}
dose_response_fr <- make_dose_response("fr", fr_plot, panel_values = c(0, pMIC_fixed))

fr_comparisons <- make_pairwise_data(
  dose_response_fr,
  list(
    "fr = 0 versus fr = 0.3" = c("fr = 0", "fr = 0.3"),
    "fr = 0 versus fr = 0.5" = c("fr = 0", "fr = 0.5")
  )
)

plot_dose_response_2d(
  fr_comparisons,
  title = "Dose-response curves by resistant fraction",
  group_title = "Resistant fraction",
  colours = fr_colours,
  file_name = "dose_response_varying_fr.png"
)
```

# Expected illness risk

For each uncertainty draw, the mean across consumption events gives the
expected illness risk. The median and 95% interval across the outer draws then
describe uncertainty in that expected risk. The treatment cut-point is where
the proportion classified as more likely treatable falls to 0.5. Expected-risk
values are transformed to log10 only for plotting.

```{r expected-risk}
add_cutoff_row <- function(data, vary_by, cutoff,
                           fixed_fr = NULL, fixed_pMIC = NULL) {
  if (!is.finite(cutoff)) return(data)

  bind_rows(
    data,
    summarise_risk(vary_by, cutoff, fixed_fr, fixed_pMIC)
  ) |>
    distinct(value, .keep_all = TRUE) |>
    arrange(value)
}

# Reference conditions
risk_summary_pMIC_fr0 <- summarise_risk("pMIC", pMIC_seq, fixed_fr = 0)
risk_summary_fr_pMIC0 <- summarise_risk("fr", fr_seq, fixed_pMIC = 0)
cutoff_pMIC_fr0 <- find_cutoff(risk_summary_pMIC_fr0, "pMIC", fixed_fr = 0, tolerance = 0.001)
cutoff_fr_pMIC0 <- find_cutoff(risk_summary_fr_pMIC0, "fr", fixed_pMIC = 0, tolerance = 0.0001)

# Main conditions
risk_summary_pMIC <- summarise_risk("pMIC", pMIC_seq, fixed_fr = fr_fixed)
risk_summary_fr <- summarise_risk("fr", fr_seq, fixed_pMIC = pMIC_fixed)
cutoff_pMIC <- find_cutoff(risk_summary_pMIC, "pMIC", fixed_fr = fr_fixed, tolerance = 0.001)
cutoff_fr <- find_cutoff(risk_summary_fr, "fr", fixed_pMIC = pMIC_fixed, tolerance = 0.0001)

risk_summary_pMIC_fr0_plot <- add_cutoff_row(risk_summary_pMIC_fr0, "pMIC", cutoff_pMIC_fr0, fixed_fr = 0)
risk_summary_fr_pMIC0_plot <- add_cutoff_row(risk_summary_fr_pMIC0, "fr", cutoff_fr_pMIC0, fixed_pMIC = 0)
risk_summary_pMIC_plot <- add_cutoff_row(risk_summary_pMIC, "pMIC", cutoff_pMIC, fixed_fr = fr_fixed)
risk_summary_fr_plot <- add_cutoff_row(risk_summary_fr, "fr", cutoff_fr, fixed_pMIC = pMIC_fixed)

expected_risk_limits <- range(
  unlist(select(bind_rows(risk_summary_pMIC_fr0_plot, risk_summary_fr_pMIC0_plot,
                          risk_summary_pMIC_plot, risk_summary_fr_plot),
                lower95_expected_risk, upper95_expected_risk)), na.rm = TRUE
)
stopifnot(expected_risk_limits[1] > 0)
expected_risk_limits <- log10(expected_risk_limits)

plot_expected_risk <- function(data, panel_label, fixed_text, x_title, x_breaks,
                               cutoff = NA_real_, cutoff_text = NULL,
                               show_y = TRUE, show_legend = FALSE) {
  data$outcome <- factor(
    data$outcome,
    levels = c("Less likely treatable", "More likely treatable")
  )

  stopifnot(all(data$lower95_expected_risk > 0,
                data$median_expected_risk > 0,
                data$upper95_expected_risk > 0))

  data <- mutate(
    data,
    log10_lower95 = log10(lower95_expected_risk),
    log10_median_risk = log10(median_expected_risk),
    log10_upper95 = log10(upper95_expected_risk)
  )

  segments <- data.frame(
    x = head(data$value, -1),
    xend = tail(data$value, -1),
    y = head(data$log10_median_risk, -1),
    yend = tail(data$log10_median_risk, -1),
    outcome = head(data$outcome, -1)
  )

  subtitle_text <- paste0(panel_label, ". ", fixed_text)
  if (is.finite(cutoff)) {
    subtitle_text <- paste0(
      subtitle_text, "; treatment cut-point = ", cutoff_text
    )
  }

  p <- ggplot(data, aes(value)) +
    geom_ribbon(
      aes(
        ymin = log10_lower95,
        ymax = log10_upper95
      ),
      fill = "grey70",
      alpha = 0.65
    ) +
    geom_segment(
      data = segments,
      aes(x = x, xend = xend, y = y, yend = yend, colour = outcome),
      linewidth = 1.05,
      show.legend = show_legend
    ) +
    scale_colour_manual(
      name = "Treatment outcome",
      values = c(
        "Less likely treatable" = "red",
        "More likely treatable" = "black"
      ),
      breaks = c("Less likely treatable", "More likely treatable"),
      drop = FALSE
    ) +
    scale_x_continuous(breaks = x_breaks) +
    scale_y_continuous(
      limits = expected_risk_limits,
      breaks = scales::breaks_pretty(n = 6),
      expand = expansion(mult = c(0.05, 0.08))
    )

  p <- p +
    labs(
      title = NULL,
      subtitle = subtitle_text,
      x = x_title,
      y = if (show_y) expression(log[10] * "(" * P[illness] * ")") else NULL
    ) +
    theme_bw(base_size = 11) +
    theme(
      plot.title = element_blank(),
      plot.subtitle = element_text(
        face = "bold", size = 11, margin = margin(b = 6)
      ),
      panel.grid.minor = element_blank(),
      legend.position = if (show_legend) "top" else "none",
      legend.box = "horizontal"
    )

  if (is.finite(cutoff)) {
    cutoff_data <- data[which.min(abs(data$value - cutoff)), ]

    p <- p +
      geom_point(
        data = cutoff_data,
        aes(x = value, y = log10_median_risk),
        inherit.aes = FALSE,
        shape = 21,
        size = 2.6,
        stroke = 0.8,
        colour = "black",
        fill = "white",
        show.legend = FALSE
      )
  }

  p
}

format_cutoff <- function(cutoff, digits, suffix = "") {
  if (!is.finite(cutoff)) return(NULL)

  paste0(
    formatC(cutoff, format = "f", digits = digits),
    suffix
  )
}

plot_A <- plot_expected_risk(
  risk_summary_pMIC_fr0_plot,
  panel_label = "A",
  fixed_text = "Resistant fraction fixed at 0",
  x_title = "Residual doxycycline concentration (pMIC)",
  x_breaks = pMIC_breaks,
  cutoff = cutoff_pMIC_fr0,
  cutoff_text = format_cutoff(cutoff_pMIC_fr0, 2, "% MIC"),
  show_legend = TRUE
)

plot_B <- plot_expected_risk(
  risk_summary_fr_pMIC0_plot,
  panel_label = "B",
  fixed_text = "pMIC fixed at 0% MIC",
  x_title = "Resistant fraction",
  x_breaks = fr_breaks,
  cutoff = cutoff_fr_pMIC0,
  cutoff_text = format_cutoff(cutoff_fr_pMIC0, 3),
  show_y = FALSE
)

plot_C <- plot_expected_risk(
  risk_summary_pMIC_plot,
  panel_label = "C",
  fixed_text = paste0("Resistant fraction fixed at ", fr_fixed),
  x_title = "Residual doxycycline concentration (pMIC)",
  x_breaks = pMIC_breaks,
  cutoff = cutoff_pMIC,
  cutoff_text = format_cutoff(cutoff_pMIC, 2, "% MIC")
)

plot_D <- plot_expected_risk(
  risk_summary_fr_plot,
  panel_label = "D",
  fixed_text = paste0("pMIC fixed at ", pMIC_fixed, "% MIC"),
  x_title = "Resistant fraction",
  x_breaks = fr_breaks,
  cutoff = cutoff_fr,
  cutoff_text = format_cutoff(cutoff_fr, 3),
  show_y = FALSE
)

column_title_pMIC <- ggplot() +
  annotate(
    "text", x = 0, y = 0.5,
    label = "Varying residual doxycycline concentration",
    hjust = 0, fontface = "bold", size = 4.5
  ) +
  xlim(0, 1) + ylim(0, 1) + theme_void()

column_title_fr <- ggplot() +
  annotate(
    "text", x = 0, y = 0.5,
    label = "Varying resistant fraction",
    hjust = 0, fontface = "bold", size = 4.5
  ) +
  xlim(0, 1) + ylim(0, 1) + theme_void()

expected_risk_four_panel <- patchwork::wrap_plots(
  A = plot_A,
  B = plot_B,
  C = plot_C,
  D = plot_D,
  E = column_title_pMIC,
  F = column_title_fr,
  G = patchwork::guide_area(),
  design = "
  GG
  EF
  AB
  CD
  ",
  heights = c(0.10, 0.07, 1, 1),
  guides = "collect"
) +
  patchwork::plot_annotation(
    title = "Expected illness risk under antibiotic and resistance scenarios",
    caption = paste(
      "Lines show log10-transformed median expected risk; ribbons show the",
      "corresponding 95% intervals from exposure and dose-response uncertainty."
    ),
    theme = theme(
      plot.title = element_text(face = "bold", size = 16),
      plot.caption = element_text(hjust = 1)
    )
  )

save_plot(
  expected_risk_four_panel,
  "expected_risk_four_panel.png",
  width = 13,
  height = 10
)
```

# Illness-risk CDF

Within each uncertainty draw, the empirical CDF describes variability among
consumption events. The median and pointwise 95% interval of these CDFs across
the outer draws show uncertainty around the variability distribution.

```{r illness-risk-cdf}
cdf_scenarios <- data.frame(
  scenario_id = c(
    "Baseline",
    "Antibiotic only",
    "Resistance only",
    "Combined"
  ),
  scenario = c(
    "Baseline (fr = 0; pMIC = 0% MIC)",
    paste0("Antibiotic only (fr = 0; pMIC = ",
           pMIC_fixed, "% MIC)"),
    paste0("Resistance only (fr = ", fr_fixed,
           "; pMIC = 0% MIC)"),
    paste0("Antibiotic and resistance (fr = ", fr_fixed,
           "; pMIC = ", pMIC_fixed, "% MIC)")
  ),
  fr = c(0, 0, fr_fixed, fr_fixed),
  pMIC = c(0, pMIC_fixed, 0, pMIC_fixed)
)

cdf_results <- pmap(cdf_scenarios[c("fr", "pMIC")], run_scenario)
log_risk <- map(cdf_results, ~log10(pmax(.x$risk, .Machine$double.xmin)))

log_risk_grid <- seq(
  min(unlist(log_risk)),
  max(unlist(log_risk)),
  length.out = 300
)

cdf_data <- map2_dfr(log_risk, seq_along(log_risk), function(x, k) {
  cdf_matrix <- vapply(
    seq_len(n_outer),
    function(j) ecdf(x[, j])(log_risk_grid),
    numeric(length(log_risk_grid))
  )

  q <- t(apply(
    cdf_matrix, 1, quantile,
    probs = c(0.025, 0.50, 0.975), names = FALSE
  ))

  data.frame(
    log10_risk = log_risk_grid,
    lower95 = q[, 1],
    median = q[, 2],
    upper95 = q[, 3],
    scenario_id = cdf_scenarios$scenario_id[k],
    scenario = cdf_scenarios$scenario[k]
  )
}) |>
  mutate(scenario = factor(scenario, levels = cdf_scenarios$scenario))

cdf_colours <- setNames(
  c("#333333", "#CC79A7", "#D55E00", "#0072B2"),
  cdf_scenarios$scenario
)

comparison_sets <- list(
  "A. Baseline vs antibiotic exposure" = c("Baseline", "Antibiotic only"),
  "B. Baseline vs resistance only" = c("Baseline", "Resistance only"),
  "C. Baseline vs combined antibiotic exposure and resistance" =
    c("Baseline", "Combined")
)

comparison_cdf_data <- imap_dfr(comparison_sets, function(scenarios, name) {
  filter(cdf_data, scenario_id %in% scenarios) |>
    mutate(comparison = name)
}) |>
  mutate(comparison = factor(comparison, levels = names(comparison_sets)))

overlap_note <- data.frame(
  comparison = factor(
    "B. Baseline vs resistance only",
    levels = names(comparison_sets)
  ),
  x = quantile(log_risk_grid, 0.25),
  y = 0.12,
  label = "Curves overlap completely"
)

illness_risk_cdf_comparisons <- ggplot(
  comparison_cdf_data,
  aes(x = log10_risk, colour = scenario, fill = scenario)
) +
  geom_ribbon(
    aes(ymin = lower95, ymax = upper95),
    alpha = 0.22,
    colour = NA
  ) +
  geom_line(aes(y = median), linewidth = 0.9) +
  geom_text(
    data = overlap_note,
    aes(x = x, y = y, label = label),
    inherit.aes = FALSE,
    size = 3.2
  ) +
  facet_wrap(~comparison, nrow = 1) +
  scale_colour_manual(name = "Scenario", values = cdf_colours) +
  scale_fill_manual(name = "Scenario", values = cdf_colours) +
  scale_y_continuous(
    limits = c(0, 1),
    breaks = seq(0, 1, by = 0.25)
  ) +
  labs(
    title = "Illness-risk variability with uncertainty: scenario comparisons",
    x = expression(log[10] * " illness probability per consumption event"),
    y = "Cumulative probability",
    caption = paste(
      "Lines show median variability CDFs; shaded areas show",
      "pointwise 95% modelled uncertainty bands."
    )
  ) +
  theme_classic(base_size = 12) +
  theme(
    plot.title = element_text(face = "bold"),
    strip.text = element_text(face = "bold"),
    legend.position = "bottom",
    legend.box = "horizontal"
  )

save_plot(
  illness_risk_cdf_comparisons,
  "illness_risk_cdf_scenario_comparisons.png",
  width = 15,
  height = 6
)
```

# Conditional Delta P_treat across the dose range

This analysis uses the fixed dose grid from the conditional dose-response
analysis. At each dose, Delta P_treat is calculated for the paired Emax-EC50
uncertainty draws. Exposure variability is not included in these figures.
Positive values indicate that a more-treatable outcome is more likely; values
at or below zero indicate that a less-treatable outcome is more likely.

```{r delta-dose-functions}
make_delta_curve <- function(fr, pMIC, group, panel) {
  q <- apply(run_dose_grid(fr, pMIC)$delta, 2, quantile, probs = c(0.025, 0.50, 0.975), names = FALSE)

  data.frame(
    dose = dose_grid,
    lower95 = q[1, ],
    median_delta = q[2, ],
    upper95 = q[3, ],
    group = group,
    panel = panel
  )
}

plot_delta_curve <- function(data, title, legend_title, colours, file_name) {
  p <- ggplot(data, aes(dose, median_delta, colour = group, fill = group)) +
    geom_ribbon(
      aes(ymin = lower95, ymax = upper95),
      alpha = 0.40,
      colour = NA,
      show.legend = FALSE
    ) +
    geom_line(linewidth = 1.05) +
    geom_hline(
      yintercept = 0,
      colour = "black",
      linetype = "dotted",
      linewidth = 0.35
    ) +
    facet_wrap(~panel, nrow = 1) +
    scale_x_log10() +
    scale_y_continuous(labels = scales::label_scientific()) +
    scale_colour_manual(
      name = legend_title,
      values = colours,
      breaks = names(colours)
    ) +
    scale_fill_manual(values = colours, breaks = names(colours)) +
    labs(
      title = title,
      x = "Pathogenic dose (MPN per consumption event)",
      y = expression(Delta * P[treat]),
      caption = expression(
        "Lines show medians; shaded areas show pointwise 95% uncertainty " *
        "intervals from paired " * E[max] * "-" * EC[50] * " simulations."
      )
    ) +
    guides(fill = "none") +
    theme_bw(base_size = 11) +
    theme(
      plot.title = element_text(face = "bold"),
      strip.text = element_text(face = "bold"),
      panel.grid.minor = element_blank(),
      legend.position = "top"
    )

  save_plot(p, file_name, width = 11, height = 6)
}
```

## Delta P_treat by residual doxycycline concentration

```{r delta-dose-pmic}
delta_curve_pMIC <- expand_grid(fr = c(0, fr_fixed), pMIC = pMIC_plot) |>
  mutate(
    group = paste0(pMIC, "% MIC"),
    panel = paste0("Resistant fraction = ", fr)
  ) |>
  pmap_dfr(make_delta_curve)

plot_delta_curve(
  delta_curve_pMIC,
  title = expression(
    Delta * P[treat] * " across dose by residual doxycycline concentration"
  ),
  legend_title = "Residual doxycycline concentration (% MIC)",
  colours = pMIC_colours,
  file_name = "delta_p_treat_across_dose_by_pMIC.png"
)
```

## Delta P_treat by resistant fraction

```{r delta-dose-fr}
delta_curve_fr <- expand_grid(pMIC = c(0, pMIC_fixed), fr = fr_plot) |>
  transmute(
    fr,
    pMIC,
    group = paste0("fr = ", fr),
    panel = paste0("pMIC = ", pMIC, "% MIC")
  ) |>
  pmap_dfr(make_delta_curve)

plot_delta_curve(
  delta_curve_fr,
  title = expression(Delta * P[treat] * " across dose by resistant fraction"),
  legend_title = "Resistant fraction",
  colours = fr_colours,
  file_name = "delta_p_treat_across_dose_by_fr.png"
)
```
